---
title: "Employee Performance Analysis"
author: "Alexander Wanstrath"
date: "2025-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Load Packages

```{r}
# Load packages
library(tidyverse)
library(stargazer)
library(ggthemes)
library(haven)
library(skimr)
library(labelled)
library(infer)
library(janitor)
library(lubridate)
library(corrr)
library(ISOweek)
library(GGally)
```

# Load Data

```{r}
## --- Load Data ---

# Load Stata files (place them in a folder called data/)
attitude_labeled <- read_dta("data/attitude_labeled.dta")
endperiod_outcomes_labeled <- read_dta("data/endperiod_outcomes_labeled.dta")
performance_labeled <- read_dta("data/performance_labeled.dta")
summary_volunteer_labeled <- read_dta("data/summary_volunteer_labeled.dta")
wage_new_labeled <- read_dta("data/wage_new_labeled.dta")

# Check if Tables are tibble
is_tibble(attitude_labeled)
is_tibble(endperiod_outcomes_labeled)
is_tibble(performance_labeled)
is_tibble(summary_volunteer_labeled)
is_tibble(wage_new_labeled)
```

# Data Cleaning

## Attitude

### Pre-Analysis

```{r}
glimpse(attitude_labeled)
skim(attitude_labeled) 
stargazer(as.data.frame(attitude_labeled), type = "text")
```

### Data Manipulation

```{r}
# Save labels
labels_attitude_labeled <- var_label(attitude_labeled)

## Change values into negative values in column negative
attitude_labeled <- attitude_labeled %>%
  mutate(negative = if_else(negative > 0, -negative, negative))

# restore labels
var_label(attitude_labeled) <- labels_attitude_labeled

# delete on same personid and year_week
attitude_labeled <- attitude_labeled %>%
  distinct(personid, year_week, .keep_all=TRUE)

### Check if Dataset is clean
skim(attitude_labeled)


```

## Endperiod

### Pre-Analysis

```{r}
glimpse(endperiod_outcomes_labeled)
skim(endperiod_outcomes_labeled) #  --> n_missing 2 for cost of commute
stargazer(as.data.frame(endperiod_outcomes_labeled), type = "text")
```

### Data Manipulation

```{r}
# Save labels
labels_endperiod_outcomes_labeled <- var_label(endperiod_outcomes_labeled)

### only numeric values - change into character

endperiod_outcomes_labeled <- endperiod_outcomes_labeled %>%
  mutate(
    costofcommute = as.character(costofcommute), # convert to character for string cleaning
    costofcommute = if_else(is.na(costofcommute), NA_character_, costofcommute),# replace NA with NA
    costofcommute = na_if(trimws(costofcommute), "-"), # replace - with NA - trimws - removes leading whitespace
    costofcommute = gsub(',','.', costofcommute), # replace ',' with '.'
    costofcommute = parse_number(costofcommute) # extract numbers
  )

### Change to factor and different labels
endperiod_outcomes_labeled <- endperiod_outcomes_labeled %>% 
  mutate(promote_switch = factor(
    x = promote_switch, levels = c(0, 1), 
    labels = c("not promoted", "promoted")
  )) 

endperiod_outcomes_labeled <- endperiod_outcomes_labeled %>% 
  mutate(quitjob = factor(
    x = quitjob, levels = c(0, 1), 
    labels = c("stayed", "quit")
  )) 

# delete duplicates on same personid
endperiod_outcomes_labeled <- endperiod_outcomes_labeled %>%
  distinct(personid, .keep_all=TRUE)

# restore labels
var_label(endperiod_outcomes_labeled) <- labels_endperiod_outcomes_labeled

### Check if Dataset is clean
skim(endperiod_outcomes_labeled)
```

## Performance

### Pre-Analysis

```{r}
glimpse(performance_labeled)
skim(performance_labeled) 
stargazer(as.data.frame(performance_labeled), type = "text")
```

### Data Manipulation

```{r}
# Save labels
labels_performance_labeled <- var_label(performance_labeled)

#Year Week - different format
performance_labeled <- performance_labeled %>% 
  mutate(year_week = gsub("[^0-9]", "",year_week)) # keep only numbers

# empty rows in column perform1 and phonecall
# phonecallraw - has some NAs
# logphonecall, logcallpersec, logcalllength and logcall_dayworked - has some empty cells and NAs

# Date - different format
performance_labeled <- performance_labeled %>%
  mutate(
    date = as.Date(parse_date_time(date, orders = c("ymd", "mdy", "dmy", 
                                                    "Ymd HMS", "Ymd HM", "Ymd H", 
                                                    "mdy HMS", "dmy HMS"))))

performance_labeled %>%
  filter(is.na(date)) %>%
  distinct(date) # Check whether my code has not found some entries -> found all

# perform1, phonecall, logphonecall --> "," into "."
performance_labeled <- performance_labeled %>%
  mutate(across(
    c(logdaysworked, phonecall, logphonecall, 
      logcallpersec, logcalllength, logcall_dayworked), # do the following for all theses tables
    ~ gsub(",", ".", .) # replace ',' with '.'
  ))

# Standardization of "-"
performance_labeled <- performance_labeled %>%
  mutate(across(where(is.character), ~gsub("[−–—‒-﹣－]", "-", .))) # replace all the different - with - for all characters

# Convert empty cells into NAs
performance_labeled <- performance_labeled %>%
  mutate(across(where(is.character), ~na_if(trimws(.x), ""))) # only on character columns

# Convert "NA" into NAs
performance_labeled <- performance_labeled %>%
  mutate(across(where(is.character), ~na_if(trimws(.x), "NA"))) # only on character columns

# Convert "-" into NAs
performance_labeled <- performance_labeled %>%
  mutate(across(where(is.character), ~na_if(trimws(.x), "-"))) # only on character columns

# mutate columns into numeric
performance_labeled <- performance_labeled %>% 
  mutate(across(
    c(perform1, phonecall, phonecallraw, logphonecall, logcallpersec, logcalllength, logcall_dayworked, logdaysworked),
    as.numeric
  ))

# mutate homethatweek into factor
performance_labeled <- performance_labeled %>% 
  mutate(homethatweek = factor(
    x = homethatweek, levels = c(0, 1), labels = c("not home", "home")
  )) 

# remove duplicates on same personid, year_week
performance_labeled <- performance_labeled %>%
  distinct(personid, year_week, .keep_all=TRUE)

# restore labels
var_label(performance_labeled) <- labels_performance_labeled

### Check if Dataset is clean
skim(performance_labeled)
```

## Volunteer

### Pre-Analysis

```{r}
glimpse(summary_volunteer_labeled)
skim(summary_volunteer_labeled) 
stargazer(as.data.frame(summary_volunteer_labeled), type = "text")
```

### Data Manipulation

```{r}
### Save labels
labels_summary_volunteer_labeled <- var_label(summary_volunteer_labeled)

### Check raw values before recoding
unique(summary_volunteer_labeled$children)
unique(summary_volunteer_labeled$bedroom)

### Change children into numeric values only (text values are recoded, not coerced to NA)
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(
    children = tolower(trimws(as.character(children))),
    children = case_when(
      children %in% c("no", "n", "false", "0") ~ 0,
      children %in% c("yes", "y", "true", "1") ~ 1,
      TRUE ~ suppressWarnings(as.numeric(children))
    )
  )

### Change bedroom into numeric values only (text values are recoded, not coerced to NA)
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(
    bedroom = tolower(trimws(as.character(bedroom))),
    bedroom = case_when(
      bedroom %in% c("no", "n", "false", "0") ~ 0,
      bedroom %in% c("yes", "y", "true", "1") ~ 1,
      TRUE ~ suppressWarnings(as.numeric(bedroom))
    )
  )

### Change Commute only numeric values 
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(commute = parse_number(as.character(commute))) # extract numbers

### Change Gender to male or female
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(
    gender = tolower(trimws(gender)),   # everything small + w/o spaces in between
    gender = case_when(
      gender %in% c("m", "male", "man") ~ "male",
      gender %in% c("f", "female", "fem", "w", "woman") ~ "female"
    )
  )

### Change Married to married or not married
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(
    married = tolower(trimws(married)),   # normalize case + spaces
    married = case_when(
      married %in% c("yes", "y", "married") ~ "married",
      married %in% c("no", "n", "not married", "single") ~ "not married"
    )
  )

### Change High_edc to high education or no high education
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(
    high_educ = tolower(trimws(high_educ)),
    high_educ = case_when(
      high_educ %in% c("yes", "y") ~ "high education",
      high_educ %in% c("no", "n")  ~ "no high education",
      TRUE ~ NA_character_
    )
  )

## Age = 1 is a data entry error. We found that people who have a tenure of 41 and 51 months, are between the ages 22 and 34.
## Therefore, we will change the values where age is 1, to the median age of 29.
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(age = if_else(age == 1, 29, age)) # change age 1 to 29

#Remove decimal numbers for columns age and tenure. Didn't apply to commute as it might be important for the analysis.
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(age = as.integer(age), tenure = as.integer(tenure)) 

# If age is 22 and tenure is 51, change the age to 28, the average age of employees with around 51 months of tenure.
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(age = if_else(age == 22 & tenure == 51, 28L, age))

# If age is 20 and tenure is 150, change the age to 35, the maximum age found in the data.
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  mutate(age = if_else(tenure == 150 & age == 20, 35L, age))

# Change Children into Factor
summary_volunteer_labeled <- summary_volunteer_labeled %>% 
  mutate(children = factor(
    x = children,
    levels = c(0,1),
    labels = c("no children", "children")
  ))

# Change Bedroom into Factor  
summary_volunteer_labeled <- summary_volunteer_labeled %>% 
  mutate(bedroom = factor(
    x = bedroom,
    levels = c(0,1),
    labels = c("no bedroom", "bedroom")
  ))

# Change gender into Factor
summary_volunteer_labeled <- summary_volunteer_labeled %>% 
  mutate(gender = factor(
    x = gender, 
    levels = c("female", "male")
  ))

# Change married into Factor
summary_volunteer_labeled <- summary_volunteer_labeled %>% 
  mutate(married = factor(
    x = married, 
    levels = c("not married", "married")
  ))

# Change high_educ into Factor
summary_volunteer_labeled <- summary_volunteer_labeled %>% 
  mutate(high_educ = factor(
    x = high_educ, 
    levels = c("no high education", "high education")
  ))

# Change Personid into numeric
summary_volunteer_labeled <- summary_volunteer_labeled %>% 
  mutate(personid = as.numeric(personid))

# delete duplicates on same personid
summary_volunteer_labeled <- summary_volunteer_labeled %>%
  distinct(personid, .keep_all=TRUE)

nrow(summary_volunteer_labeled) # check --> correct 135 rows

### restore labels
var_label(summary_volunteer_labeled) <- labels_summary_volunteer_labeled

### Check if Dataset is clean
skim(summary_volunteer_labeled) # 4 NAs
```

## Wage

### Pre-Analysis

```{r}
glimpse(wage_new_labeled)
skim(wage_new_labeled) 
stargazer(as.data.frame(wage_new_labeled), type = "text")
```

### Data Manipulation

```{r}
# Save labels
labels_wage_new_labeled <- var_label(wage_new_labeled)

### only numeric values - change into character
wage_new_labeled <- wage_new_labeled %>%
  mutate(
    basewage = as.character(basewage),         # make sure it's text
    basewage = trimws(basewage),               # remove leading/trailing spaces
    basewage = gsub("[^0-9,]", "", basewage),  # keep only digits and commas
    basewage = gsub(",00$", "", basewage),     # remove trailing ,00
    basewage = gsub(",", "", basewage),        # remove other commas
    basewage = as.numeric(basewage)            # back to numeric
  )

wage_new_labeled <- wage_new_labeled %>%
  mutate(
    grosswage = as.character(grosswage), # convert to character for string cleaning
    grosswage = if_else(is.na(grosswage), NA_character_, grosswage),# replace NA with NA
    grosswage = na_if(trimws(grosswage), "-"), # replace - with NA - trimws - removes leading whitespace
    grosswage = gsub(',','.', grosswage), # replace ',' with '.'
    grosswage = as.numeric(gsub("[^0-9.-]", "", grosswage)) # extract numbers
  )

wage_new_labeled <- wage_new_labeled %>%
  mutate(
    bonustotal = as.character(bonustotal), #create as character - for the following operations
    bonustotal = if_else(is.na(bonustotal), NA_character_, bonustotal),# replace NA with NA
    bonustotal = na_if(trimws(bonustotal), "-"), # replace - with NA - trimws - removes leading whitespace
    bonustotal = gsub(',','.', bonustotal), # replace ',' with '.'
    bonustotal = as.numeric(gsub("[^0-9.-]", "", bonustotal)) # extract numbers
  )

wage_new_labeled <- wage_new_labeled %>%
  mutate(
    personid = as.character(personid),#create as character - for the following operations
    personid = trimws(personid), # remove leading and trailing whitespace
    personid = gsub("[^0-9]", "", personid), # keep only numeric characters
    personid = sub("^0+", "", personid), # remove leading zeros
    personid = as.numeric(personid) # convert to numeric
  )

wage_new_labeled <- wage_new_labeled %>%
  mutate(
    wage_month = gsub("[^0-9]", "", wage_month),                     # keep only digits
    wage_month = substr(wage_month, nchar(wage_month)-5, nchar(wage_month)), # last 6 digits
    wage_month = if_else(nchar(wage_month) == 5,     # fix 5-digit -> add 0
                         paste0(substr(wage_month,1,4),"0",substr(wage_month,5,5)),
                         wage_month),
    wage_month = if_else(nchar(wage_month) == 4, paste0(wage_month, "01"), false = wage_month), # YYYY -> YYYY01
    wage_month = paste0(substr(wage_month,1,4), "-", substr(wage_month,5,6)) # format YYYY-MM
  )

# Change column name wage_month into year_month for merge
wage_new_labeled <- wage_new_labeled %>%
  rename(year_month = wage_month)

# remove duplicates on same personid, yearmonth
wage_new_labeled <- wage_new_labeled %>%
  distinct(personid, year_month, .keep_all=TRUE)

## Filter if there are NAs in 2 of 3 Columns
wage_new_labeled <- wage_new_labeled %>% filter(!is.na(basewage) & !is.na(grosswage) | !is.na(bonustotal) & !is.na(grosswage) | !is.na(basewage) & !is.na(bonustotal))

# Calculate Base Wage
wage_new_labeled <- wage_new_labeled %>% mutate(basewage = ifelse(is.na(basewage), grosswage-bonustotal, basewage))

# Calculate Bonus
wage_new_labeled <- wage_new_labeled %>% mutate(bonustotal = ifelse(is.na(bonustotal), grosswage-basewage, bonustotal))

# Recalculate Gross Wage
wage_new_labeled <- wage_new_labeled %>% mutate(grosswage = ifelse(grosswage != bonustotal+basewage, bonustotal+basewage, grosswage))

# Columns as numeric
wage_new_labeled <- wage_new_labeled %>%
  mutate(across(c(basewage, grosswage, bonustotal), as.numeric))

### Check if Dataset is clean
skim(wage_new_labeled) 
```

# Merge Tables

## Preparation

We must create Year_Month Column for all tables to be able to merge them.

Current Tables:

-   "Performance" has Year_Week –\> we must create a column for Year_month

-   "Volunteer" has no Date –\> No need for adjustments

-   "Wage" has Wage_Month –\> rename it to Year_month

-   "Endperiod" has no Date –\> No need for adjustments

-   "Attitude" has Year week –\> we have to create Year_month

```{r}
# Performance create new Column - Year_Month
performance_labeled_monthly <- performance_labeled %>%
  mutate(
    year_month = format(
        ISOweek::ISOweek2date( # Type of Date
          paste0(substr(as.character(year_week), 1, 4), # Slice Year
                 "-W", # insert "-"
                 substr(as.character(year_week), 5, 6), # Slice Week
                 "-1") # first day of the Week
        ),
        "%Y-%m"
    )
  )

# Attitude create new Column - Year_Month
attitude_labeled_cleaned_monthly <- attitude_labeled %>% 
  mutate(
    year_month = format(
        ISOweek::ISOweek2date( # Type of Date
          paste0(substr(as.character(year_week), 1, 4), # Slice Year
                 "-W", # insert "-"
                 substr(as.character(year_week), 5, 6), # Slice Week
                 "-1") # first day of the Week
        ),
        "%Y-%m" # always day one
    )
  )
```

## Merge

1.  Merge tables that have no date
    -   Tables: Endperiod and Volunteer

```{r}
# Merge Tables that have no date
Volunteer_endperiod_merged <- merge(summary_volunteer_labeled, endperiod_outcomes_labeled, by = "personid", all = TRUE)
```

2.  Merge 1. with one of the tables that have Year_month
    -   Table: Wage

```{r}
Volunteer_endperiod_wage_merged <- merge(wage_new_labeled, Volunteer_endperiod_merged, by = "personid", all = TRUE)
```

3.  Merge 2. with one of the other tables that have Year_month
    -   Table: Performance

```{r}
Volunteer_endperiod_wage_performance_merged <- merge(performance_labeled_monthly, Volunteer_endperiod_wage_merged, by = c("personid", "year_month"), all = TRUE)
```

4.  Merge last table with 3.
    -   Table: Attitude

```{r}
all_tables <- merge(attitude_labeled_cleaned_monthly, Volunteer_endperiod_wage_performance_merged, by = c("personid", "year_month", "year_week"), all = TRUE)
```

## Check Skim

```{r}
skim(all_tables)
skim(attitude_labeled_cleaned_monthly) # --> same statistics
skim(performance_labeled_monthly) # --> same statistics
skim(endperiod_outcomes_labeled) # --> same statistics
skim(summary_volunteer_labeled) # --> same statistics
skim(wage_new_labeled) # --> different stats
```

# Analysis

## Helper Functions and Person-Level Table

Person-level characteristics (gender, education, promotion, quitting, tenure, age, commute cost) are constant within a person but repeated in every week of the panel. Testing them on the panel would count each person many times and overstate significance. These tests therefore use `person_level`, which has one row per employee.

```{r}
# Decision and p-value formatting for t-tests
h0_decision <- function(test, alpha = 0.05) {
  if (test$p.value < alpha) "Reject H0" else "Fail to reject H0"
}
p_fmt <- function(test) format.pval(test$p.value, digits = 3, eps = 0.001)

# One row per employee
person_level <- all_tables %>%
  group_by(personid) %>%
  summarise(
    avg_performance = mean(perform1, na.rm = TRUE),
    across(c(age, tenure, commute, costofcommute), ~ first(na.omit(.x))),
    across(c(gender, married, high_educ, children, bedroom, promote_switch, quitjob),
           ~ first(na.omit(.x))),
    .groups = "drop"
  ) %>%
  mutate(avg_performance = if_else(is.nan(avg_performance), NA_real_, avg_performance))

nrow(person_level)
```

## Means Table

Create a table with the mean for each Column per Personid

```{r}
Mean_Personid <- all_tables %>%
group_by(personid) %>%
  summarise(
    across(
      where(is.numeric) & !any_of("personid"),
      \(x) { if (all(is.na(x))) NA_real_ else mean(x, na.rm = TRUE) },
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )
```

## Explore Dataset

```{r}
skim(all_tables)
```

## Summary Statistics

```{r}
stargazer(data.frame(all_tables), type = "text"
          ,iqr = FALSE, median = TRUE, no.space = TRUE, title = "Summary")
```

## Correlation Matrix

```{r}
cor_matrix_all <- all_tables %>% 
  select(where(is.numeric), -personid)%>% # 
  correlate(diagonal = 1)

# Format table for a professional presentation
matrix_all <- cor_matrix_all %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation

```

## Workforce Profile

### Workers total

```{r}
plot_workers <- all_tables %>%
  summarise(total_persons = n_distinct(personid)) %>% 
  ggplot(aes(x = "All Workers", y = total_persons)) +  
  geom_col(fill = "#4C78A8", color = "white") +          
  labs(
    x = "",
    y = "Number of Unique Persons",
    title = "Total Number of Workers"
  ) +
  geom_text(aes(label = total_persons), vjust = -0, size = 4) + # Number on top of Bar
  theme_minimal() +
    theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title.y = element_text(margin = margin(r = 10))
  )

plot_workers

```

### Age

```{r}
# Histogram Age
plot_age <- all_tables %>%
  filter(!is.na(age)) %>% 
  group_by(personid) %>% 
  summarise(age = first(age)) %>% 
  ggplot(aes(x = age)) + 
  geom_histogram(binwidth = 1, fill = "#4C78A8", color = "white") +
  labs(x = "Age", y = "Count", title = "Age Distribution") + labs(x="Age") +
  theme_minimal()

plot_age
```

```{r}
Age_Profile <- all_tables %>% 
  group_by(personid) %>% 
  summarize( min_age = min(age),
             max_age = max(age),
             avg_age = mean(age))

Age_Profile
```

### Gender

```{r}
plot_gender <- all_tables %>%
  filter(!is.na(gender)) %>%
  group_by(personid) %>% 
  summarise(gender = first(gender)) %>% 
  ggplot(aes(x = gender, fill = gender)) + 
  geom_bar(color = "white") +
  geom_text(stat = "count", aes(label = after_stat(count)),vjust = 0 , size = 4) +  # <- Text on top of the bar
  scale_fill_manual(values = c( female ="#4C78A8", male = "#A6CEE3")) +
  labs(x = "Gender", y = "Count", title = "Gender Distribution") + labs(x="Gender") +
  theme_minimal() +
    theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title.y = element_text(margin = margin(r = 10))
  )


plot_gender
```

### Education

```{r}
plot_education <- all_tables %>%
  filter(!is.na(high_educ)) %>%
  group_by(personid) %>% 
  summarise(high_educ = first(high_educ)) %>% 
  ggplot(aes(x = high_educ, fill = high_educ)) + 
  geom_bar(color = "white") +
  geom_text(stat = "count", aes(label = after_stat(count)),vjust = 0 , size = 4) +  # <- Text on top of the bar
  scale_fill_manual(values = c( "no high education" ="#4C78A8","high education"= "#A6CEE3")) +
  labs(x = "Education", y = "Count", title = "Education Distribution") + labs(x="Education") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title.y = element_text(margin = margin(r = 10))
  )


plot_education
```

### Marital Status

```{r}
plot_married <- all_tables %>%
  filter(!is.na(married)) %>%
  group_by(personid) %>%
  summarise(married = first(married), .groups = "drop") %>%
  ggplot(aes(x = married, fill = married)) +
  geom_bar(color = "white") +
  geom_text(
    stat = "count",
    aes(label = after_stat(count)),
    vjust = -0.3,   # slightly above the bar
    size = 4
  ) +
  scale_fill_manual(
    values = c(
      "not married" = "#4C78A8",
      "married" = "#A6CEE3"
    )
  ) +
  labs(
    x = "Marital Status",
    y = "Count",
    title = "Marital Status Distribution"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title.y = element_text(margin = margin(r = 10))
  )

plot_married

```

## Attitude

### Summary Statistics

```{r}
stargazer(data.frame(attitude_labeled), type = "text"
          ,iqr = FALSE, median = TRUE, no.space = TRUE, title = "Summary Attitude")
```

### Correlation Table

```{r}
cor_matrix_attitude <- all_tables %>% 
  select(exhaustion, negative, positive) %>% 
  correlate(diagonal = 1)

# Format table for a professional presentation
cor_matrix_attitude %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation
```

### Attitude - Histogram

```{r}
# Histogram Exhaustion
ggplot( data = all_tables, aes(x = exhaustion)) + geom_histogram( fill = "#4C78A8", color = "white") + labs(x="Exhaustion") +
  theme_minimal()

# Histogram negative
ggplot( data = all_tables, aes(x = negative)) + geom_histogram(fill = "#4C78A8", color = "white") + labs(x="Negative") +
  theme_minimal()

# Histogram positive
ggplot( data = all_tables, aes(x = positive)) + geom_histogram(fill = "#4C78A8", color = "white") + labs(x="Positive") +
  theme_minimal()
```

### Attitude - Scatterplot

```{r}
# Scatterplot Exhaustion and Negative
plot.scatter_attitude_negative <-  ggplot(data = all_tables
                                          , aes(x = exhaustion, y = negative)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Exhaustion"
         , y = "Negative"
         , title = "Relationship between Exhaustion and Negative") +
  theme(title = element_text(size = 8)) +
  theme_minimal()

# Add regression Line
plot.scatter_attitude_negative + stat_smooth(method = lm, colour="blue")
```

```{r}
# Scatterplot Exhaustion and positive
plot.scatter_attitude_positive <-  ggplot(data = all_tables
                                          , aes(x = exhaustion, y = positive)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Exhaustion"
         , y = "Positive"
         , title = "Relationship between Exhaustion and Positive") +
  theme(title = element_text(size = 8))

# Add regression Line
plot.scatter_attitude_positive + stat_smooth(method = lm, colour="blue") +
  theme_minimal()
```

```{r}
# Scatterplot negative and positive
plot.scatter_attitude_negative_positive <-  ggplot(data = all_tables
                                                   , aes(x = negative, y = positive)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Negative"
         , y = "Positive"
         , title = "Relationship between negative and Positive") +
  theme(title = element_text(size = 8)) +
  theme_minimal()

# Add regression Line
plot.scatter_attitude_negative_positive + stat_smooth(method = lm, colour="blue")
```

## Performance

```{r}
summary(all_tables$perform1)
```

### Variable Distribution

```{r}
# Histogram Perform1
ggplot( data = all_tables, aes(x = perform1)) + geom_histogram(fill = "#4C78A8", color = "white") +
  labs( x = "Performance") +
  theme_minimal()

# check with Boxplot
ggplot(data = performance_labeled, aes(x = perform1)) + 
  geom_boxplot(fill = "#4C78A8") + 
  labs( x = "Performance") +
  scale_y_discrete() + 
  theme_minimal()
# some positive outliers, a lot of negative ones

```

### Summary Table

```{r}
# Create Summary Table
collapse_factor <- function(x) {
  x <- x[!is.na(x)]                     # remove NAs
  if (length(x) == 0) {                 # if all missing
    return(factor(NA, levels = levels(x)))
  } else {
    return(factor(x[1], levels = levels(x)))  # keep factor type
  }
}

# Main pipeline
performance_groups_all <- all_tables %>%
  filter(!is.na(gender)) %>%            # keep only rows with valid gender
  group_by(personid) %>%
  summarise(
    avg_performance = mean(perform1, na.rm = TRUE),
    across(where(is.factor), collapse_factor),  # take one value per factor column
    .groups = "drop"
  ) %>%
  mutate(
    perf_group = case_when(
      avg_performance >= 0.5  ~ "Top performer",
      avg_performance <= -0.5 ~ "Low performer",
      TRUE                    ~ "Average performer"
    )
  )
```

### Perform / Phonecallraw

#### Variable Relationships

```{r}
#Variable Relationship
# Scatterplot Perform and Phonecallraw
plot.scatter_performance_phonecallraw <-  ggplot(data = all_tables
                                    , aes(x = perform1, y = phonecallraw)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Performance"
         , y = "Phonecalls"
         , title = "Relationship between Performance and Phonecalls") +
  theme(title = element_text(size = 8)) +
  theme_minimal()

# Add regression Line
plot.scatter_performance_phonecallraw + stat_smooth(method = lm, colour="blue")
```

#### Correlation Table

```{r}

cor_matrix_performance_Phonecallraw <- all_tables %>% 
  select(perform1, phonecallraw)%>% # 
  correlate(diagonal = 1)

# Format table for a professional presentation
cor_matrix_performance_Phonecallraw %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation
```

The correlation between performance and phone calls is almost perfect (r = 0.95). This most likely reflects how the performance measure is constructed, since it is largely based on call volume, rather than an independent finding.

### Performance / Homethatweek

#### Distribution

```{r}
all_tables %>%
  filter(!is.na(homethatweek), !is.na(perform1)) %>% 
  ggplot(aes(x = homethatweek, y = perform1, fill = homethatweek)) + 
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(title = "Performance by Homethatweek"
         , x = "homethatweek"
         , y = "Performance") +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = "Performance" is not influenced by "homethatweek"

```{r}
# t.test if the mean performance is better at home or not at home
tt_home <- t.test(perform1 ~ homethatweek, data = all_tables, alternative = "less")
tt_home
```

–\> `r h0_decision(tt_home)` (p = `r p_fmt(tt_home)`)

Mean weekly performance: not home `r round(tt_home$estimate[1], 3)`, home `r round(tt_home$estimate[2], 3)`.

### Phonecalls / Homethatweek

```{r}
all_tables %>%
  filter(!is.na(homethatweek), !is.na(phonecall)) %>% 
  ggplot(aes(x = homethatweek, y = phonecall, fill = homethatweek)) + 
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(title = "Phonecall by Homethatweek"
         , x = "homethatweek"
         , y = "Phonecall") +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = "Phonecall" is not influenced by "homethatweek"

```{r}
tt_calls <- t.test(phonecall ~ homethatweek, data = all_tables, alternative = "less")
tt_calls
```

–\> `r h0_decision(tt_calls)` (p = `r p_fmt(tt_calls)`)

Mean weekly calls: not home `r round(tt_calls$estimate[1], 3)`, home `r round(tt_calls$estimate[2], 3)`.

### Performance / Logdaysworked

#### Correlation Table

```{r}

cor_matrix_performance_logdaysworked <- all_tables %>% 
  select(perform1, logdaysworked)%>% # 
  correlate(diagonal = 1)

# Format table for a professional presentation
cor_matrix_performance_logdaysworked %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation

```

The correlation table shows a moderate positive correlation between performance and logdaysworked (r = 0.47).

#### Variable Relationship - Scatter plot

```{r}
# Visualization of the Correlation
plot.scatter_performance_logdaysworked <-  ggplot(data = all_tables
                                                 , aes(x = perform1, y = logdaysworked)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Performance"
         , y = "Logdaysworked"
         , title = "Relationship between Performance and Days worked") +
  theme(title = element_text(size = 8)) +
  theme_minimal()

# Add regression Line
plot.scatter_performance_logdaysworked + stat_smooth(method = lm, colour="blue")
```

The scatter plot shows a positive relationship between performance and logdaysworked.

### Performance / Exhaustion

#### Correlation Table

```{r}
cor_matrix_performance_exhaustion <- all_tables %>% 
  select(perform1, exhaustion, negative, positive)%>% # 
  correlate(diagonal = 1)

# Format table for a professional presentation
cor_matrix_performance_exhaustion %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation

```

There is no real correlation between Performance and Exhaustion.

### Performance / Base Wage

#### Variable Relationship - Scatter plot

```{r}
avg_df <- all_tables %>%
  group_by(year_month) %>%
  summarise(
    avg_base_wage = mean(basewage, na.rm = TRUE),
    avg_performance = mean(perform1, na.rm = TRUE)
  )


ggplot(avg_df, aes(x = avg_base_wage, y = avg_performance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", colour="blue") +
  labs(
    title = "Average Performance vs Average Base Wage",
    subtitle = "By Year-Month",
    x = "Average Base Wage (€)",
    y = "Average Performance"
  ) +
  theme_minimal()
```

The scatter plot compares monthly averages. At the individual level, the correlation between base wage and performance is close to zero (r = -0.02), so the monthly pattern should not be read as a link between individual pay and performance.

### Performance / Gross Wage

#### Variable Relationship - Scatter plot

```{r}
avg_df <- all_tables %>%
  group_by(year_month) %>%
  summarise(
    avg_gross_wage = mean(grosswage, na.rm = TRUE),
    avg_performance = mean(perform1, na.rm = TRUE)
  )


ggplot(avg_df, aes(x = avg_gross_wage, y = avg_performance)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", colour="blue") +
  labs(
    title = "Average Performance vs Average Gross Wage",
    subtitle = "By Year-Month",
    x = "Average Gross Wage (€)",
    y = "Average Performance"
  ) +
  theme_minimal()

```

As with base wage, the plot shows monthly averages and does not reflect the individual level relationship.

#### Correlation Matrix

```{r}
cor_matrix_2 <- all_tables %>% 
  select(perform1, grosswage) %>% 
  correlate(diagonal = 1)

# Format table
cor_matrix_2 %>%
  rearrange() %>% 
  shave() %>% 
  fashion(decimals = 3)

```

At the individual level there is practically no correlation between performance and gross wage (r = 0.01).

### Performance / Bonus

#### Correlation

```{r}
cor_matrix_performance_bonus <- all_tables %>% 
  select(perform1, basewage, grosswage, bonustotal)%>% # 
  correlate(diagonal = 1)

# Format table for a professional presentation
cor_matrix_performance_bonus %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation
```

The correlation table shows a weak positive correlation between performance and bonus (r = 0.25), while base and gross wage are unrelated to performance.

#### Variable Relationship - Scatter plot

```{r}
# Scatterplot Performance - Bonus
plot.scatter_performance_bonus <-  ggplot(data = all_tables
                                          , aes(x = perform1, y = bonustotal)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Performance"
         , y = "Bonus"
         , title = "Relationship between Performance and Bonus") +
  theme(title = element_text(size = 8)) +
  theme_minimal()

# Add regression Line
plot.scatter_performance_bonus + stat_smooth(method = lm, colour="blue")
```

The scatter plot confirms the weak positive relationship between bonus and performance.

### Performance / Age

#### Correlation Table

```{r}
cor_matrix_performance_age <- all_tables %>% 
  select(perform1, age)%>% # 
  correlate(diagonal = 1)

# Format table for a professional presentation
cor_matrix_performance_age %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation
```

–\> Only a very weak correlation between performance and age (r = 0.08).

#### Variable Relationship - Scatter plot

```{r}
plot.scatter_performance_age_mean <-  ggplot(data = Mean_Personid
                                             , aes(x = mean_perform1, y = mean_age)) +
  geom_point(size = 1) +
  theme_few() +
  labs(  x = "Performance"
         , y = "age"
         , title = "Relationship between Performance and age") +
  theme(title = element_text(size = 8)) +
  theme_minimal()

# Add regression Line
plot.scatter_performance_age_mean + stat_smooth(method = lm, colour="blue")
```

**Home Performance vs. Age**

```{r}
# Bar chart Performance by Age (home weeks only)
# 1) Prepare data: filter home only, valid age & performance
df_home <- all_tables %>%
  filter(homethatweek == "home",
         !is.na(age),
         !is.na(perform1))

df_summary <- df_home %>%
  mutate(age_group = age) %>%
  group_by(age_group) %>%
  summarise(
    n = dplyr::n(),
    mean_performance = mean(perform1, na.rm = TRUE),
    sd = sd(perform1, na.rm = TRUE),
    se = sd / sqrt(n),
    ci95 = 1.96 * se,
    .groups = "drop"
  )


ggplot(df_summary, aes(x = age_group, y = mean_performance)) +
  geom_col(fill = "steelblue", color = "white") +
  labs(
    title = "Average Performance by Age Group (Home Only)",
    x = "Age Group",
    y = "Average Performance (perform1)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle =45,hjust=1))
```

The chart shows average performance in work-from-home weeks for each age.

### Performance / Quit

#### Variable Relationship - Box plot

```{r}
all_tables %>% 
filter(!is.na(quitjob), !is.na(perform1)) %>%
  ggplot(aes(x = quitjob, y = perform1, fill = quitjob)) +
  geom_boxplot() +
    stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(
    title = "Performance by Quit Status",
    x = "Quit Status",
    y = "Performance Score"
  ) +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = Employees who quit and employees who stayed have the same average performance

```{r}
tt_quit <- t.test(avg_performance ~ quitjob, data = person_level)
tt_quit
```

–\> `r h0_decision(tt_quit)` (p = `r p_fmt(tt_quit)`)

Mean performance: stayed `r round(tt_quit$estimate[1], 3)`, quit `r round(tt_quit$estimate[2], 3)`.

**Bar chart showing number of people by Quit status and performance group**

```{r}
performance_groups_all %>%
  ggplot(aes(x = perf_group, fill = quitjob)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of Performance Groups by Quit Status",
    x = "Performance Group",
    y = "Number of Employees"
  ) +
  theme_minimal()
```

### Performance / Promotion

#### Variable Relationship - Box plot

```{r}
all_tables %>%
  filter(!is.na(promote_switch), !is.na(perform1)) %>%
  ggplot(aes(x = promote_switch, y = perform1, fill = promote_switch)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(
    title = "Performance by Promotion Status",
    x = "Promotion Status",
    y = "Performance Score"
  ) +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = Promoted and not promoted employees have the same average performance

```{r}
tt_promo <- t.test(avg_performance ~ promote_switch, data = person_level)
tt_promo
```

–\> `r h0_decision(tt_promo)` (p = `r p_fmt(tt_promo)`)

Mean performance: not promoted `r round(tt_promo$estimate[1], 3)`, promoted `r round(tt_promo$estimate[2], 3)`.

**Bar chart showing number of people by promotion status and performance group**

```{r}
performance_groups_all %>%
  ggplot(aes(x = perf_group, fill = promote_switch)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of Performance Groups by Promotion Status",
    x = "Performance Group",
    y = "Number of Employees"
  ) +
  theme_minimal()
```

### Performance / Education

#### Variable Relationship - Box plot

```{r}
all_tables %>% 
  filter(!is.na(high_educ), !is.na(perform1)) %>% 
  group_by(personid) %>% 
  ggplot(aes(x = high_educ, y = perform1, fill = high_educ)) + 
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(
    title = "Performance by Education",
    x = "Education",
    y = "Performance Score"
  ) +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = Employees with and without high education have the same average performance

```{r}
tt_educ <- t.test(avg_performance ~ high_educ, data = person_level)
tt_educ
```

–\> `r h0_decision(tt_educ)` (p = `r p_fmt(tt_educ)`)

Mean performance: no high education `r round(tt_educ$estimate[1], 3)`, high education `r round(tt_educ$estimate[2], 3)`.

**Bar chart showing number of people by Education and performance group**

```{r}
performance_groups_all %>%
  ggplot(aes(x = perf_group, fill = high_educ)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of Performance Groups by Education",
    x = "Performance Group",
    y = "Number of Employees"
  ) +
  theme_minimal()
```

### Performance / Tenure

#### Variable Relationship - Scatter plot

```{r}
all_tables %>%
  filter(!is.na(tenure), !is.na(perform1)) %>%
  ggplot(aes(x = tenure, y = perform1)) +
  geom_point(alpha = 0.4) +        # scatter points
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # regression line
  labs(
    title = "Relationship between Tenure and Performance",
    x = "Tenure (months)",
    y = "Performance Score"
  ) +
  theme_minimal()
```

The scatter plot shows a very weak positive relationship between tenure and performance (r = 0.11).

### Performance / Gender

#### Variable Relationship - Box plot

```{r}
all_tables %>%
  filter(!is.na(gender), !is.na(perform1)) %>%
  ggplot(aes(x = gender, y = perform1, fill = gender)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(
    title = "Performance by Gender",
    x = "Gender",
    y = "Performance Score"
  ) +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = Female and male employees have the same average performance

```{r}
tt_gender <- t.test(avg_performance ~ gender, data = person_level)
tt_gender
```

–\> `r h0_decision(tt_gender)` (p = `r p_fmt(tt_gender)`)

Mean performance: female `r round(tt_gender$estimate[1], 3)`, male `r round(tt_gender$estimate[2], 3)`.

#### Performance / Gender

**Bar chart showing number of people by gender and performance group**

```{r}
performance_groups_all %>%
  ggplot(aes(x = perf_group, fill = gender)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of Performance Groups by Gender",
    x = "Performance Group",
    y = "Number of Employees"
  ) +
  theme_minimal()
```

The chart compares the number of female and male employees in each performance group.

### Performance Overview

#### Correlation Table

```{r}
# Select numerical variables
Performance_numeric_var <- all_tables %>% 
  select(perform1, phonecallraw, logdaysworked, exhaustion, basewage, grosswage, bonustotal, age, tenure)

# Correlation Table between numerical variables
correlation_matrix_performance <- correlate(Performance_numeric_var, diagonal = 1)

# Print Table
correlation_matrix_performance %>%
  rearrange() %>%              # rearrange by correlations
  shave() %>%                  # Shave off the upper triangle for a clean result
  fashion(decimals = 3)        # Clean presentation

# Visualize the correlation matrix
ggcorr(Performance_numeric_var,
       label = TRUE,
       method = c("everything", "pearson"),
       low = "blue", mid = "white", high = "red",
       layout.exp = 2,          # increases the distance from text to matrix
       label_size = 3,
       hjust = 0.9, vjust = 0.9)
```

The correlation matrix shows:

-   a very strong correlation between performance and phone calls (r = 0.95), which reflects how performance is measured
-   a moderate positive correlation with days worked (r = 0.47)
-   a weak positive correlation with bonus (r = 0.25)
-   no meaningful correlation with exhaustion, base wage, gross wage, age or tenure (all |r| ≤ 0.11)

## Endperiod

### Cost of Commute

#### Variable Distribution - Histogram

```{r}
ggplot(all_tables, aes(x = costofcommute)) +
  geom_histogram(
    binwidth = 5,                      # adjust bins to your data scale
    fill = "#4C78A8",
    color = "white",                    # clean bin borders
  ) +
  labs(
    title = "Distribution of Monthly Commute Cost",
    x = "Monthly Commute Cost",
    y = "Number of Employees"
  ) +
  theme_minimal(base_size = 13)
```

#### Variable Distribution - Boxplot

```{r}
ggplot(data = all_tables, aes(x = costofcommute)) + 
  geom_boxplot( fill ="#4C78A8") + 
  scale_y_discrete() +
  theme_minimal()

```

### Cost of Commute / Commute (Time)

#### Variable Relationships - Scatter plot

```{r}
all_tables %>%
  filter(!is.na(commute), !is.na(costofcommute)) %>%
  ggplot(aes(x = commute, y = costofcommute)) +
  geom_point(alpha = 0.4) +        # scatter points
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # regression line
  labs(
    title = "Relationship between Commute time and Cost of commute",
    x = "commute (minutes)",
    y = "Cost of commute"
  ) +
  theme_minimal()
```

Scatter plot shows that there is a correlation between Cost of commute and commute (minutes).

### Cost of Commute / Quit Job

#### Descriptive Statistics

```{r}
all_tables %>% 
  filter(!is.na(quitjob), !is.na(costofcommute)) %>% 
  group_by(personid) %>% 
  ggplot(aes(x = quitjob, y = costofcommute, fill = quitjob)) + 
  geom_boxplot() +
  labs(
    title = "Cost of Commute by Quit Status",
    x = "Quit Status",
    y = "Cost of Commute"
  ) +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = Employees who quit and employees who stayed have the same average commute cost

```{r}
tt_commute <- t.test(costofcommute ~ quitjob, data = person_level)
tt_commute
```

–\> `r h0_decision(tt_commute)` (p = `r p_fmt(tt_commute)`)

Mean monthly commute cost: stayed `r round(tt_commute$estimate[1], 2)`, quit `r round(tt_commute$estimate[2], 2)`.

### Cost of Commute / Exhaustion

#### Variable Relationships - Scatter plot

```{r}
ggplot(all_tables, aes(x = costofcommute, y = exhaustion)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Cost of Commute vs Exhaustion", x = "Cost of Commute", y = "Exhaustion") +
  theme_minimal()
```

The scatter plot shows a slight negative correlation.

## Volunteer

### Tenure

#### Tenure / Married

##### Variable Relationship - Box plot

```{r}
ggplot(summary_volunteer_labeled %>% drop_na(married, tenure),
       aes(x = married, y = tenure, fill = married)) +
  geom_boxplot(outlier.alpha = 0.25) +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(title = "Tenure by Marital Status", x = "Marital status", y = "Tenure (months)") +
  theme_minimal()
```

##### T.Test

–\> Hypothesis: H0 = Married and not married employees have the same average tenure

```{r}
tt_tenure_married <- t.test(tenure ~ married, data = person_level)
tt_tenure_married
```

–\> `r h0_decision(tt_tenure_married)` (p = `r p_fmt(tt_tenure_married)`)

Mean tenure in months: not married `r round(tt_tenure_married$estimate[1], 1)`, married `r round(tt_tenure_married$estimate[2], 1)`.

#### Age / Married

##### Variable Relationship - Box plot

```{r}
ggplot(summary_volunteer_labeled %>% drop_na(married, age),
       aes(x = married, y = age, fill = married)) +
  geom_boxplot(outlier.alpha = 0.25) +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(title = "Age by Marital Status", x = "Marital status", y = "Age (years)") +
  theme_minimal()
```

##### T.Test

–\> Hypothesis: H0 = Married and not married employees have the same average age

```{r}
tt_age_married <- t.test(age ~ married, data = person_level)
tt_age_married
```

–\> `r h0_decision(tt_age_married)` (p = `r p_fmt(tt_age_married)`)

Mean age: not married `r round(tt_age_married$estimate[1], 1)`, married `r round(tt_age_married$estimate[2], 1)`.

#### Tenure / Age

##### Variable Relationship - Scatter plot

```{r}
ggplot(all_tables, aes(x = tenure, y = age)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Tenure and Age", x = "Tenure (months)", y = "Age") +
  theme_minimal()
```

The scatter plot shows a very high correlation between Tenure and Age.

Since tenure and age are closely related, differences in tenure by marital status may largely reflect differences in age.

#### Tenure / Exhaustion

##### Variable Relationship - Scatter plot

```{r}
ggplot(all_tables, aes(x = tenure, y = exhaustion)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Tenure vs Exhaustion", x = "Tenure (months)", y = "Exhaustion") +
  theme_minimal()
```

This Plot shows that there only is a slight correlation between Exhaustion and Tenure.

#### Tenure / Promotion

##### Variable Relationship - Box plot

```{r}
ggplot(person_level %>% drop_na(promote_switch, tenure),aes(x = promote_switch, y = tenure, fill = promote_switch)) +
  geom_boxplot(outlier.alpha = 0.25) +
  stat_summary(fun = mean, geom = "point", size = 3, shape = 23) +
  labs(title = "Tenure by Promotion Status", x = "Promotion Status", y = "Tenure (months)") +
  theme_minimal()
```

#### T.Test

–\> Hypothesis: H0 = Promoted and not promoted employees have the same average tenure

```{r}
tt_tenure_promo <- t.test(tenure ~ promote_switch, data = person_level)
tt_tenure_promo
```

–\> `r h0_decision(tt_tenure_promo)` (p = `r p_fmt(tt_tenure_promo)`)

Mean tenure in months: not promoted `r round(tt_tenure_promo$estimate[1], 1)`, promoted `r round(tt_tenure_promo$estimate[2], 1)`.
